### 2025-02-06 sections/experimental.tex
Reformatted the main tex file for customization, and changed the sections to match requirements

\subsection{\label{sec:radio}The Radio Lightcurve}

\subsection{\label{sec:optical}The Optical Lightcurve}