### 2025-02-06 uca_radio_eclipse_jurpa.tex

\documentclass[%
 aip,
% jmp,
% bmf,
% sd,
% rsi,
cp,  % Conference Proceedings
 amsmath,amssymb,%nobibnotes,
% preprint,%
 reprint,%
%author-year,%
%author-numerical,%
]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%% Loads a Times-like font. You can also load
%% {newtxtext,newtxtmath}, but not {times}, 
%% {txfonts} nor {mathtpm} as these packages
%% are obsolete and have been known to cause problems.
\usepackage{mathptmx}

\begin{document}

\title{Radio Telescope Observations Of The April 8, 2024 Total Solar Eclipse At 1420MHz.}% Force line breaks with \\

\author{August Childress} % Write as First name Surname
\email[Corresponding author: ]{achildress@cub.uca.edu}
\affiliation{Department of Physics, Astronomy, and Engineering, University of Central Arkansas, Conway, AR 72035, USA}
\author{Blayne Griffin}%
\email{bgriffin@cub.uca.edu.}
\affiliation{Department of Physics, Astronomy, and Engineering, University of Central Arkansas, Conway, AR 72035, USA}
\author{Jeremy Lusk}
\email{jlusk@uca.edu}
\affiliation{Department of Physics, Astronomy, and Engineering, University of Central Arkansas, Conway, AR 72035, USA}

\date{\today} % It is always \today, today, but any date may be explicitly specified
              % Not printed for conference proceedings

\begin{abstract}
\input{sections/abstract}
\end{abstract}

\maketitle

\section{\label{sec:introduction}Introduction}
\input{sections/introduction}

\section{\label{sec:experimental}Experimental Investigations}
\input{sections/experimental}

\section{\label{sec:theoretical}Theoretical Studies}
\input{sections/theoretical}

\section{\label{sec:conclusion}Conclusion}
\input{sections/conclusion}

\begin{acknowledgments}
\input{sections/acknowledgements}
\end{acknowledgments}

\nocite{*}
\bibliography{refs}% Produces the bibliography via BibTeX.

\end{document}
%
% ****** End of file aipsamp.tex ******