### 2025-03-06 sections/acknowledgements.tex
Removed Technical Specs from acknowledgements

\noindent Special thanks to these people for providing us the data necessary for our project:

% Come back and indent and make parentheses affiliations
\begin{enumerate}
    \item Scott Austin, Ph.D. (Director of Astronomical Facilities, Associate Professor of Astronomy and Physics, Department of Physics, Astronomy, and Engineering, University of Central Arkansas) 
    \item Todd Abel, Ph.D. (UCA STEM Institute Co-Director, Associate Professor of Mathematics, Secondary Mathematics Education Program Coordinator)
    \item Shane Ayotte. (UCA Mathematics Student)
\end{enumerate}