### 2025-03-07 sections/conclusion.tex
Cleaned up Sections and edited Conclusion I cleaned up general stuff in Observations and Theoretical, including an explanation as to the variance in radio vs theoretical curves. I also edited the conclusion to match these changes.

Based on our comparison of data from a 2.3 m radio telescope, a broad-spectrum light sensor, an optical livestream, and theoretical data from \texttt{Stellarium}\cite{zotti_simulated_2020} and JPL Horizons \cite{nasa_jpl_solar_system_dynamics_group_jpl_nodate}, our hypothesis that the radio solar disk is larger than the optical solar disk is supported.
This analysis found that the best-fit curve has an apparent solar radius of $R_{\mathrm{1420}} = 1.27 R_{\odot}$, lying between the results of other studies \cite{messerotti_radio_2000} and \cite{leung_solar_2022}.
In future work, we will compare these and other results to examine potential sources of uncertainty in our measurements and apparent variance in data.
We will also explore different modeling techniques to further characterize the radio emissions from the Sun.