### 2025-03-14 sections/acknowledgements.tex
Updating grants section

\noindent Special thanks to Scott Austin, Ph.D. (Director of Astronomical Facilities, Associate Professor of Astronomy and Physics, Department of Physics, Astronomy, and Engineering, University of Central Arkansas), Todd Abel, Ph.D. (UCA STEM Institute Co-Director, Associate Professor of Mathematics, Secondary Mathematics Education Program Coordinator), and Shane Ayotte. (UCA Mathematics Student)

This research would not have been possible without the support of state and federal research grants. Funding to purchase the radio telescope came from the Arkansas Space Grant Consortium (ASGC) Statewide Equipment Award, grant number GR013133-22-UCA. Funding to support student research and travel to present our results came from the Arkansas Space Grant Consortium (ASGC) Student Intensive Training award, grant number 242040-25.UCA.SIT.Lusk, a subaward of NASA award number 80NSSC20M0106.