### 2025-03-14 sections/acknowledgements.tex
Updaing the award numbers and adding github links

\noindent Special thanks to Scott Austin, Ph.D. (Director of Astronomical Facilities, Associate Professor of Astronomy and Physics, Department of Physics, Astronomy, and Engineering, University of Central Arkansas), Todd Abel, Ph.D. (UCA STEM Institute Co-Director, Associate Professor of Mathematics, Secondary Mathematics Education Program Coordinator), and Shane Ayotte (UCA Mathematics Student).

Our research team practices open-source science. All data and analysis code can be accessed at \url{https://github.com/UCA-Open-Astronomy-Lab/total_solar_eclipse}. The repository for this paper can be found at \url{https://github.com/UCA-Open-Astronomy-Lab/jurpa_eclipse_paper}

This research would not have been possible without the support of state and federal research grants. Funding to purchase the radio telescope came from the Arkansas Space Grant Consortium (ASGC) Statewide Equipment Award, matched by ASGC state match 20-ARMF-02. Funding to support student research and travel to present our results came from the ASGC Student Intensive Training award. These ASGC awards are subawards of NASA award number 80NSSC20M0106.