### 2025-03-15 sections/acknowledgements.tex

\noindent Special thanks to Scott Austin, Ph.D. (Director of Astronomical Facilities, Associate Professor of Astronomy and Physics, Department of Physics, Astronomy, and Engineering, University of Central Arkansas), Todd Abel, Ph.D. (UCA STEM Institute Co-Director, Associate Professor of Mathematics, Secondary Mathematics Education Program Coordinator), and Shane Ayotte (UCA Mathematics Student).

Our research team practices open-source science. All data and analysis code, along with the source of this paper can be accessed at \url{https://github.com/UCA-Open-Astronomy-Lab}

This research would not have been possible without the support of state and federal grants. Funding to purchase the radio telescope came from the Arkansas Space Grant Consortium (ASGC) Statewide Equipment Award, matched by ASGC state match 20-ARMF-02. Funding to support student research and travel to present our results came from the ASGC Student Intensive Training award. These ASGC awards are subawards of NASA award 80NSSC20M0106.