### 2025-03-15 sections/conclusion.tex

Based on our comparison of data from a 2.3 m radio telescope, a broad-spectrum light sensor, an optical livestream, and theoretical data from \texttt{Stellarium} and JPL Horizons, our hypothesis that the radio solar disk is larger than the optical solar disk is supported.
This analysis found that the best-fit curve has an apparent solar radius of $R_{\mathrm{1420}} = 1.27 R_{\odot}$, falling between the results of other studies \cite{messerotti_radio_2000} and \cite{leung_solar_2022}.
In future work, we hope to examine the radio lightcurve in detail with the goal of mapping radio emission from the sun's surface as features are eclipsed by the moon's edge moving across the face of the sun, as done in \cite{messerotti_radio_2000}.