### 2025-05-12 response.tex
Adding response file

\documentclass{article}
\begin{document}

We are greatful to the reviewer for their insightful suggestions to improve our manuscript.
Included below are summaries of the reviewer's suggested changes, and our modifications to the manuscript.

\section{Savitsky-Golay filtering}

From the reviewer:

\begin{quote}
The authors also use a well-tested filtering method for accounting for interference.
I am curious, though, why was this method chosen over others? I suspect it may be that it is straightforward, which reduces the possibility of introducing more errors and biases, but it is not clear.
\end{quote}
\noindent We have added the following:

\begin{quote}
  \input{sections/savitsky_golay.tex}
\end{quote}

\end{document}