### 2025-05-12 sections/model_slope.tex

In future work, we hope to examine the radio lightcurve in detail with the goal of analyzing the variance seen between our model slope and the recorded radio data. In a previous study by \cite{messerotti_radio_2000} the slopes also differ, but not to the degree seen in our data.
In \cite{messerotti_radio_2000}, that variance is used to map radio emission from the sun's surface as features are eclipsed by the moon's edge moving across the face of the sun, something we hope to replicate in a future study.