### 2025-05-14 sections/acknowledgements.tex
Making acknowledgements section shorter

\noindent Special thanks to Scott Austin, Ph.D. (Associate Professor of Astronomy and Physics, UCA), Todd Abel, Ph.D. (UCA STEM Institute Co-Director, Associate Professor of Mathematics), and Shane Ayotte (UCA Mathematics Student).
All data and analysis code, along with the source of this paper can be accessed at \url{https://github.com/UCA-Open-Astronomy-Lab}.
Funding for this research came from an Arkansas Space Grant Consortium (ASGC) Statewide Equipment Award, and an ASGC Student Intensive Training award. These ASGC awards are subawards of NASA award 80NSSC20M0106.